---
title: "Explore Intro Scenario"
author: "Cat Pien, Liz Stebbins, Flowwest"
format: html
editor: visual
---

# Example Portfolio Model Results

## List brief description of alternatives included




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r message = FALSE, warning = FALSE}
library(winterRunDSM)
library(tidyverse)
library(plotly)
library(ggplot2)
library(zoo)
library(kableExtra)

theme_plots <- 
  theme_bw() + 
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 13),
        legend.position = "top")

colors <- c("#96C8A2", "#006A4E")
colors2 <- c("darkgoldenrod2", "coral4")
```


```{r, warning = FALSE}
# baseline
baseline_seeds <- winterRunDSM::winter_run_model(scenario = NULL, 
                                                 mode = "seed",
                                                 seeds = NULL, 
                                                 ..params = winterRunDSM::wr_sdm_baseline_params)

baseline_results <- winterRunDSM::winter_run_model(mode = "simulate", 
                                                   ..params = winterRunDSM::wr_sdm_baseline_params,
                                                   seeds = baseline_seeds)

# test out scenario 
alt_seeds <- winterRunDSM::winter_run_model(scenario = NULL, 
                                            mode = "seed",
                                            seeds = NULL, 
                                            ..params = winterRunDSM::wr_sdm_scen_intro_params)

alt_results <- winterRunDSM::winter_run_model(mode = "simulate", 
                                              ..params = winterRunDSM::wr_sdm_scen_intro_params,
                                              seeds = alt_seeds)
```

## Summary Table of all the metrics 


## Abundance 

### Spawners
```{r, out.width = '100%'}
spawners <- baseline_results$spawners |> 
  as.data.frame() |> 
  mutate(watershed = winterRunDSM::watershed_labels,
         scenario = "baseline") |> 
  bind_rows(alt_results$spawners |> 
              as.data.frame() |> 
              mutate(watershed = winterRunDSM::watershed_labels,
                     scenario = "alt")) |> 
  pivot_longer(`1`:`20`,
               names_to = "sim_year",
               values_to = "spawners") |> 
  filter(watershed %in% c("Upper Sacramento River", "Battle Creek")) |> 
  mutate(sim_year = as.numeric(sim_year)) 

mean_spawners <- spawners |> 
  group_by(scenario, watershed) |> 
  summarize(mean_spawners = round(mean(spawners))) |> 
  arrange(watershed)

plotly::ggplotly(ggplot(spawners) + 
  geom_line(aes(sim_year, spawners, color = scenario)) +
  geom_point(aes(sim_year, spawners, color = scenario, shape = scenario), size = 3) +
    facet_wrap(~watershed, nrow = 2) + 
  scale_color_manual(values = colors)+
  theme_plots)

# 
# * 20-year mean spawners for baseline: `r mean_spawners$mean_spawners[mean_spawners$scenario=="baseline"]`
# * 20-year mean spawners for alt: `r mean_spawners$mean_spawners[mean_spawners$scenario=="alt"]`
# * 20-year mean spawners for baseline: `r mean_spawners$mean_spawners[mean_spawners$scenario=="baseline"]`
# * 20-year mean spawners for alt: `r mean_spawners$mean_spawners[mean_spawners$scenario=="alt"]`
```


```{r}
kable(mean_spawners)
```


### Natural origin WR returning to spawn

```{r, echo = FALSE, out.width = '100%'}
spawn_prop <- alt_results$proportion_natural_at_spawning[c("Upper Sacramento River", "Battle Creek"),] |> 
  as.data.frame() |> 
  mutate(watershed = c("Upper Sacramento River", "Battle Creek")) |> 
  pivot_longer(`1`:`20`,
               names_to = "sim_year",
               values_to = "prop_natural") |> 
  mutate(sim_year = as.numeric(sim_year))

returns <- left_join(spawners, spawn_prop) |> 
  mutate(nat_returns = spawners * prop_natural,
         hatchery_returns = spawners - nat_returns)

mean_natural_return <- returns |>
  group_by(scenario, watershed) |>
  summarize(mean_natural_return = round(mean(nat_returns))) |> 
  arrange(watershed)

ggplot(returns) +
  # geom_col(aes(return_sim_year, returns, fill = scenario), color = "black", position = position_dodge2()) +
  geom_line(aes(sim_year, nat_returns, color = scenario)) +
  geom_point(aes(sim_year, nat_returns, color = scenario, shape = scenario), size = 3) +
  labs(title = "Natural Origin Returning Adults") +
  facet_wrap(~watershed, nrow = 2) + 
  scale_color_manual(values = colors)+
  theme_plots


# nat_return <- baseline_results$returning_adults |> 
#   as.data.frame() |> 
#   mutate(scenario = "baseline") |> 
#   filter(origin == "natural") |> 
#   bind_rows(alt_results$returning_adults |> 
#               as.data.frame() |> 
#               mutate(scenario = "alt") |> 
#                filter(origin == "natural")) |> 
#   filter(watershed %in% c("Upper Sacramento River")) |> 
#   group_by(watershed, return_sim_year, scenario) |> 
#   summarize(returns = sum(return_total))
# 
# mean_natural_return <- nat_return |> 
#   group_by(scenario) |> 
#   summarize(mean_natural_return = round(mean(returns)))
# 
# ggplot(nat_return) + 
#   # geom_col(aes(return_sim_year, returns, fill = scenario), color = "black", position = position_dodge2()) +
#   geom_line(aes(return_sim_year, returns, color = scenario)) +
#   geom_point(aes(return_sim_year, returns, color = scenario, shape = scenario), size = 3) +
#   labs(title = "Natural Origin Returning Adults") +
#   scale_color_manual(values = colors)+
#   theme_plots
```

```{r}
kable(mean_natural_return)
```

### Hatchery origin WR returning to spawn

```{r, echo = FALSE, out.width = '100%'}
mean_hatchery_return <- returns |>
  group_by(scenario, watershed) |>
  summarize(mean_hatchery_return = round(mean(hatchery_returns))) |> 
  arrange(watershed)

ggplot(returns) +
  # geom_col(aes(return_sim_year, returns, fill = scenario), color = "black", position = position_dodge2()) +
  geom_line(aes(sim_year, hatchery_returns, color = scenario)) +
  geom_point(aes(sim_year, hatchery_returns, color = scenario, shape = scenario), size = 3) +
  labs(title = "Hatchery Origin Returning Adults") +
  facet_wrap(~watershed, nrow = 2) + 
  scale_color_manual(values = colors2)+
  theme_plots

# hatchery_return <- baseline_results$returning_adults |> 
#   as.data.frame() |> 
#   mutate(scenario = "baseline") |> 
#   filter(origin == "hatchery") |> 
#   bind_rows(alt_results$returning_adults |> 
#               as.data.frame() |> 
#               mutate(scenario = "alt") |> 
#                filter(origin == "hatchery")) |> 
#   filter(watershed %in% c("Upper Sacramento River")) |> 
#   group_by(watershed, return_sim_year, scenario) |> 
#   summarize(returns = sum(return_total))
# 
# mean_hatchery_return <- hatchery_return |> 
#   group_by(scenario) |> 
#   summarize(mean_hatchery_return = round(mean(returns)))
# 
# ggplot(nat_return) + 
#   # geom_col(aes(return_sim_year, returns, fill = scenario), color = "black", position = position_dodge2()) +
#   geom_line(aes(return_sim_year, returns, color = scenario)) +
#   geom_point(aes(return_sim_year, returns, color = scenario, shape = scenario), size = 3) +
#   labs(title = "Hatchery Origin Returning Adults") +
#   scale_color_manual(values = colors2)+
#   theme_plots
```

```{r}
kable(mean_hatchery_return)
```

### Frequency of population decline
```{r}
decline <- spawners |> 
  group_by(scenario, watershed) |> 
  mutate(decline = if_else(spawners < lag(spawners,1), 1L, 0L)) |> 
  ungroup()

total_declines <- decline |> 
  group_by(scenario, watershed) |> 
  summarize(total_declines = sum(decline, na.rm = TRUE)) |> 
  arrange(watershed)


# * Number of years of declining population for baseline: `r total_declines$total_declines[total_declines$scenario=="baseline"]`
# * Number of years of declining population for alt: `r total_declines$total_declines[total_declines$scenario=="alt"]`
```


```{r}
kable(total_declines)
```


### Frequency of catastrophic decline

```{r}
change_spawn <- spawners |> 
  group_by(scenario, watershed) |> 
  mutate(change_abund = spawners-lag(spawners, 1),
         prop_change_abund = change_abund/lag(spawners,1)) |> 
  mutate(avg_change = rollapply(change_abund, width = 3, FUN = mean, align = "right", partial = FALSE, fill = NA),
         avg_prop_change = rollapply(prop_change_abund, width = 3, FUN = mean, align = "right", partial = FALSE, fill = NA) ) |> 
  ungroup()

max_declines <- change_spawn |> 
  group_by(scenario, watershed) |> 
  summarize(max_decline = round(min(avg_prop_change, na.rm = TRUE),3)) |> 
  arrange(watershed)
```

```{r}
kable(max_declines)
```


## Productivity

### Number of juveniles emigrating from upper Sacramento River

Number of juveniles coming out of Upper Sacramento River before rear, grow, migrate, moving to other watersheds

```{r}
juvs_us <- baseline_results$spawners |>
  as.data.frame() |>
  mutate(watershed = winterRunDSM::watershed_labels,
         scenario = "baseline") |>
  bind_rows(alt_results$spawners |>
              as.data.frame() |>
              mutate(watershed = winterRunDSM::watershed_labels,
                     scenario = "alt")) |>
  pivot_longer(`1`:`20`,
               names_to = "sim_year",
               values_to = "spawners") |>
  filter(watershed %in% c("Upper Sacramento River", "Battle Creek")) |>
  mutate(sim_year = as.numeric(sim_year))


juvs_us <- baseline_results$juveniles |>
   mutate(watershed = winterRunDSM::watershed_labels,
         scenario = "baseline") |>

alt_results$juveniles |>
  filter(watershed == "Upper Sacramento River") |>
  group_by(watershed, year) |>
  summarize(total_juv = sum(juveniles))
  ungroup

ggplot(juvs_us) +
  # geom_col(aes(return_sim_year, returns, fill = scenario), color = "black", position = position_dodge2()) +
  geom_line(aes(year, total_juv, color = scenario)) +
  geom_point(aes(year, total_juv, color = scenario, shape = scenario), size = 3) +
  labs(title = "Juveniles in Upper Sac", y = "Juveniles in Upper Sac") +
  scale_color_manual(values = colors)+
  theme_plots
```


### Number of natural origin smolts surviving to reach the ocean 

Do we want to view different sizes and return months? Grouped right now.

```{r, out.width = '100%'}
juv_chipps <- baseline_results$juveniles_at_chipps |> 
  as.data.frame() |> 
  mutate(scenario = "baseline") |> 
  bind_rows(alt_results$juveniles_at_chipps |> 
              as.data.frame() |> 
              mutate(scenario = "alt")) |>  
  filter(watershed %in% c("Upper Sacramento River", "Battle Creek"),
         size %in% c("l", "vl")) |> 
  group_by(watershed, scenario, year) |> 
  summarize(jac = sum(juveniles_at_chipps)) |> 
  ungroup()

mean_jac <- juv_chipps |> 
  group_by(scenario, watershed) |> 
  summarize(mean_jac = round(mean(jac))) |> 
  arrange(watershed)

ggplot(juv_chipps) + 
  # geom_col(aes(return_sim_year, returns, fill = scenario), color = "black", position = position_dodge2()) +
  geom_line(aes(year, jac, color = scenario)) +
  geom_point(aes(year, jac, color = scenario, shape = scenario), size = 3) +
  labs(title = "Juveniles at Chipps", y = "Juveniles at Chipps") +
  facet_wrap(~watershed, nrow = 2) + 
  scale_color_manual(values = colors)+
  theme_plots
```



```{r}
kable(mean_jac)
```


### Cohort Replacement Rate
```{r, out.width = '100%'}
crr <- spawners |> 
  group_by(scenario, watershed) |> 
  mutate(crr = spawners/lag(spawners, 3),
         crr_over_one = if_else(crr > 1, 1L, 0L)) |> 
  ungroup()

crr_summary <- crr |> 
  group_by(scenario, watershed) |> 
  summarize(years_over_one = sum(crr_over_one, na.rm = TRUE),
         mean_crr = round(mean(crr, na.rm = TRUE),3)) |> 
  ungroup() |> 
  arrange(watershed)

ggplot(crr) + 
  geom_line(aes(sim_year, crr, color = scenario)) +
  geom_point(aes(sim_year, crr, color = scenario, shape = scenario), size = 3) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  labs(title = "Cohort Replacement Rate", y = "Cohort Replacement Rate") +
  facet_wrap(~watershed, nrow = 2) + 
  scale_color_manual(values = colors)+
  theme_plots
```

```{r}
kable(crr_summary)
```


## Life history diversity and fitness
Add labels to the y axis 

### pHOS
```{r, out.width = '100%'}
phos <- baseline_results$phos |> 
  as.data.frame() |> 
  mutate(watershed = winterRunDSM::watershed_labels,
         scenario = "baseline") |> 
  bind_rows(alt_results$phos |> 
              as.data.frame() |> 
              mutate(watershed = winterRunDSM::watershed_labels,
                     scenario = "alt")) |> 
  pivot_longer(`1`:`20`,
               names_to = "sim_year",
               values_to = "phos") |> 
  filter(watershed %in% c("Upper Sacramento River", "Battle Creek")) |> 
  mutate(sim_year = as.numeric(sim_year)) 
  
ggplot(phos) + 
  geom_line(aes(sim_year, phos, color = scenario)) +
  geom_point(aes(sim_year, phos, color = scenario, shape = scenario), size = 3) +
  scale_color_manual(values = colors2)+
  labs(title = "pHOS") +
  facet_wrap(~watershed, nrow = 2) + 
  theme_plots
```





### Size class diversity

```{r}

```


### Timing diversity 
```{r}

```


### Combined size-timing diversity 
```{r}
juvenile_size_ocean_entry <- baseline_results$juveniles_at_chipps |>
    group_by(year, month, watershed, size_or_age = size) |>
    summarise(value = sum(juveniles_at_chipps, na.rm = TRUE)) |>
    mutate(scenario = "baseline") |> bind_rows(
  alt_results$juveniles_at_chipps |>
    group_by(year, month, watershed, size_or_age = size) |>
    summarise(value = sum(juveniles_at_chipps, na.rm = TRUE)) |>
    mutate(scenario = "alt")) |> 
  filter(watershed %in% c("Upper Sacramento River", "Battle Creek"))

  annual_total <- juvenile_size_ocean_entry |>
    group_by(year, scenario) |>
    summarize(total_juveniles = sum(value, na.rm = T))

  shannon_di <- juvenile_size_ocean_entry |>
    group_by(year, size_or_age, scenario, month) |>
    summarize(frequency = sum(value, na.rm = T)) |>
    ungroup() |>
    left_join(annual_total) |>
    mutate(pi = frequency / total_juveniles,
           ln_pi = log(pi),
           pi_ln_pi = pi * ln_pi) |>
    group_by(year, scenario) |>
    summarize(shannon_index = -1 * sum(pi_ln_pi, na.rm = T)) |>
    ungroup()
  
  mean_shannon_di <- shannon_di |>
    group_by(scenario) |>
    summarise(avg_annual_shannon_di = mean(shannon_index))
```

```{r}
juv_size_ocean_entry <- juvenile_size_ocean_entry |> mutate(juvgp = paste0(month, size_or_age))
ggplot(juv_size_ocean_entry)+
  geom_col(aes(year, value, fill = juvgp), position = "fill", color = "black") +
  facet_wrap(~size_or_age)+ 
  theme_bw()

```

```{r}
ggplot(shannon_di) + 
  geom_line(aes(year, shannon_index, color = scenario)) +
  geom_point(aes(year, shannon_index, color = scenario, shape = scenario), size = 3) +
  scale_color_manual(values = colors2)+
  labs(title = "Shannon Diversity for Timing and Size") +
  theme_plots
  
```



```{r}

```


## Increase number of tribs supporting WR

### Number of adults spawning in tributaries

```{r}
spawners_in_tribs <- spawners  |>  filter(watershed == "Battle Creek")

mean_spawners_in_tribs <- mean_spawners |> filter(watershed == "Battle Creek")

ggplot(spawners_in_tribs) + 
  geom_line(aes(sim_year, spawners, color = scenario)) +
  geom_point(aes(sim_year, spawners, color = scenario, shape = scenario), size = 3) +
  labs(title = "Spawners in Battle Creek")+
  scale_color_manual(values = colors)+
  theme_plots
```

```{r}
kable(mean_spawners_in_tribs)
```

### Number of juveniles rearing in tributaries

```{r}

```


### Number of independent populations
```{r}
ind_pop <- spawners |> left_join(phos) |> left_join(crr) |> left_join(change_spawn) |> 
  mutate(growth_rate = change_abund/lag(spawners)) |> 
  mutate(above_500_spawners = if_else(spawners > 500, TRUE, FALSE),
         phos_less_than_5_percent = if_else(phos < .05, TRUE, FALSE),
         crr_above_1 = if_else(crr >= 1, TRUE, FALSE),
         growth_rate_above_1 = if_else(growth_rate >= 0, TRUE, FALSE),
         independent_population = if_else(above_500_spawners & phos_less_than_5_percent &
                                           growth_rate_above_1 & crr_above_1, TRUE, FALSE))
```


### Number of independent populations in historic habitats


### Number of dependent populations 

