---
title: "Explore Intro Scenario"
author: "Cat Pien, Liz Stebbins, Flowwest"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r message = FALSE, warning = FALSE}
library(winterRunDSM)
library(tidyverse)
library(plotly)
library(ggplot2)
library(zoo)

theme_plots <- 
  theme_bw() + 
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 13),
        legend.position = "top")

colors <- c("#96C8A2", "#006A4E")
colors2 <- c("darkgoldenrod2", "coral4")
```


```{r, warning = FALSE}
# baseline
baseline_seeds <- winterRunDSM::winter_run_model(scenario = NULL, 
                                                 mode = "seed",
                                                 seeds = NULL, 
                                                 ..params = winterRunDSM::wr_sdm_baseline_params)

baseline_results <- winterRunDSM::winter_run_model(mode = "simulate", 
                                                   ..params = winterRunDSM::wr_sdm_baseline_params,
                                                   seeds = baseline_seeds)

# test out scenario 
alt_seeds <- winterRunDSM::winter_run_model(scenario = NULL, 
                                            mode = "seed",
                                            seeds = NULL, 
                                            ..params = winterRunDSM::wr_sdm_scen_intro_params)

alt_results <- winterRunDSM::winter_run_model(mode = "simulate", 
                                              ..params = winterRunDSM::wr_sdm_scen_intro_params,
                                              seeds = alt_seeds)
```

## Abundance 

### Spawners
```{r, out.width = '100%'}
spawners <- baseline_results$spawners |> 
  as.data.frame() |> 
  mutate(watershed = winterRunDSM::watershed_labels,
         scenario = "baseline") |> 
  bind_rows(alt_results$spawners |> 
              as.data.frame() |> 
              mutate(watershed = winterRunDSM::watershed_labels,
                     scenario = "alt")) |> 
  pivot_longer(`1`:`20`,
               names_to = "sim_year",
               values_to = "spawners") |> 
  filter(watershed %in% c("Upper Sacramento River")) |> 
  mutate(sim_year = as.numeric(sim_year)) 

mean_spawners <- spawners |> 
  group_by(scenario) |> 
  summarize(mean_spawners = round(mean(spawners)))

plotly::ggplotly(ggplot(spawners) + 
  geom_line(aes(sim_year, spawners, color = scenario)) +
  geom_point(aes(sim_year, spawners, color = scenario, shape = scenario), size = 3) +
  scale_color_manual(values = colors)+
  theme_plots)
```

* 20-year mean spawners for baseline: `r mean_spawners$mean_spawners[mean_spawners$scenario=="baseline"]`
* 20-year mean spawners for alt: `r mean_spawners$mean_spawners[mean_spawners$scenario=="alt"]`

### Natural origin WR returning to spawn

Do we want sim_year or return_sim_year? Do we want to break these down into age classes?  

```{r, echo = FALSE, out.width = '100%'}
nat_return <- baseline_results$returning_adults |> 
  as.data.frame() |> 
  mutate(scenario = "baseline") |> 
  filter(origin == "natural") |> 
  bind_rows(alt_results$returning_adults |> 
              as.data.frame() |> 
              mutate(scenario = "alt") |> 
               filter(origin == "natural")) |> 
  filter(watershed %in% c("Upper Sacramento River")) |> 
  group_by(watershed, return_sim_year, scenario) |> 
  summarize(returns = sum(return_total))

mean_natural_return <- nat_return |> 
  group_by(scenario) |> 
  summarize(mean_natural_return = round(mean(returns)))

ggplot(nat_return) + 
  # geom_col(aes(return_sim_year, returns, fill = scenario), color = "black", position = position_dodge2()) +
  geom_line(aes(return_sim_year, returns, color = scenario)) +
  geom_point(aes(return_sim_year, returns, color = scenario, shape = scenario), size = 3) +
  labs(title = "Natural Origin Returning Adults") +
  scale_color_manual(values = colors)+
  theme_plots
```


* 20-year mean natural return for baseline: `r mean_natural_return$mean_natural_return[mean_natural_return$scenario=="baseline"]`
* 20-year mean natural return for alt: `r mean_natural_return$mean_natural_return[mean_natural_return$scenario=="alt"]`


### Hatchery origin WR returning to spawn

```{r, echo = FALSE, out.width = '100%'}
hatchery_return <- baseline_results$returning_adults |> 
  as.data.frame() |> 
  mutate(scenario = "baseline") |> 
  filter(origin == "hatchery") |> 
  bind_rows(alt_results$returning_adults |> 
              as.data.frame() |> 
              mutate(scenario = "alt") |> 
               filter(origin == "hatchery")) |> 
  filter(watershed %in% c("Upper Sacramento River")) |> 
  group_by(watershed, return_sim_year, scenario) |> 
  summarize(returns = sum(return_total))

mean_hatchery_return <- hatchery_return |> 
  group_by(scenario) |> 
  summarize(mean_hatchery_return = round(mean(returns)))

ggplot(nat_return) + 
  # geom_col(aes(return_sim_year, returns, fill = scenario), color = "black", position = position_dodge2()) +
  geom_line(aes(return_sim_year, returns, color = scenario)) +
  geom_point(aes(return_sim_year, returns, color = scenario, shape = scenario), size = 3) +
  labs(title = "Hatchery Origin Returning Adults") +
  scale_color_manual(values = colors2)+
  theme_plots
```

* 20-year mean hatchery return for baseline: `r mean_hatchery_return$mean_hatchery_return[mean_hatchery_return$scenario=="baseline"]`
* 20-year mean hatchery return for alt: `r mean_hatchery_return$mean_hatchery_return[mean_hatchery_return$scenario=="alt"]`

### Frequency of population decline
```{r}
decline <- spawners |> 
  group_by(scenario) |> 
  mutate(decline = if_else(spawners < lag(spawners,1), 1L, 0L)) |> 
  ungroup()

total_declines <- decline |> 
  group_by(scenario) |> 
  summarize(total_declines = sum(decline, na.rm = TRUE))
```

* Number of years of declining population for baseline: `r total_declines$total_declines[total_declines$scenario=="baseline"]`
* Number of years of declining population for alt: `r total_declines$total_declines[total_declines$scenario=="alt"]`

### Frequency of catastrophic decline

```{r}
change_spawn <- spawners |> 
  group_by(scenario) |> 
  mutate(change_abund = spawners-lag(spawners, 1),
         prop_change_abund = change_abund/lag(spawners,1)) |> 
  mutate(avg_change = rollapply(change_abund, width = 3, FUN = mean, align = "right", partial = FALSE, fill = NA),
         avg_prop_change = rollapply(prop_change_abund, width = 3, FUN = mean, align = "right", partial = FALSE, fill = NA) ) |> 
  ungroup()

max_declines <- change_spawn |> 
  group_by(scenario) |> 
  summarize(max_decline = round(min(avg_prop_change, na.rm = TRUE),3))
```


* Largest 3-year-avg decline for baseline: `r max_declines$max_decline[max_declines$scenario=="baseline"]`
* Largest 3-year-avg decline for alt: `r max_declines$max_decline[max_declines$scenario=="alt"]`

## Productivity

### Number of juveniles emigrating from upper Sacramento River

Is this output$juveniles or do we need to get from somewhere else? 


### Number of natural origin smolts surviving to reach the ocean 

Do we want to view different sizes and return months? Grouped right now.

```{r, out.width = '100%'}
juv_chipps <- baseline_results$juveniles_at_chipps |> 
  as.data.frame() |> 
  mutate(scenario = "baseline") |> 
  bind_rows(alt_results$juveniles_at_chipps |> 
              as.data.frame() |> 
              mutate(scenario = "alt")) |>  
  filter(watershed %in% c("Upper Sacramento River"),
         size %in% c("l", "vl")) |> 
  group_by(watershed, scenario, year) |> 
  summarize(jac = sum(juveniles_at_chipps)) |> 
  ungroup()

mean_jac <- juv_chipps |> 
  group_by(scenario) |> 
  summarize(mean_jac = round(mean(jac)))

ggplot(juv_chipps) + 
  # geom_col(aes(return_sim_year, returns, fill = scenario), color = "black", position = position_dodge2()) +
  geom_line(aes(year, jac, color = scenario)) +
  geom_point(aes(year, jac, color = scenario, shape = scenario), size = 3) +
  labs(title = "Juveniles at Chipps", y = "Juveniles at Chipps") +
  scale_color_manual(values = colors)+
  theme_plots
```

* Mean Juveniles at Chipps for baseline: `r mean_jac$mean_jac[mean_jac$scenario=="baseline"]`
* Mean Juveniles at Chipps for alt: `r mean_jac$mean_jac[mean_jac$scenario=="alt"]`

### Cohort Replacement Rate
```{r, out.width = '100%'}
crr <- spawners |> 
  group_by(scenario) |> 
  mutate(crr = spawners/lag(spawners, 3),
         crr_over_one = if_else(crr > 1, 1L, 0L)) |> 
  ungroup()

crr_summary <- crr |> 
  group_by(scenario) |> 
  summarize(years_over_one = sum(crr_over_one, na.rm = TRUE),
         mean_crr = round(mean(crr, na.rm = TRUE),3)) |> 
  ungroup()

ggplot(crr) + 
  geom_line(aes(sim_year, crr, color = scenario)) +
  geom_point(aes(sim_year, crr, color = scenario, shape = scenario), size = 3) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  labs(title = "Cohort Replacement Rate", y = "Cohort Replacement Rate") +
  scale_color_manual(values = colors)+
  theme_plots
```

* Mean CRR for baseline: `r crr_summary$mean_crr[crr_summary$scenario=="baseline"]`
* Mean CRR for alt: `r crr_summary$mean_crr[crr_summary$scenario=="alt"]`
* Years with CRR > 1 for baseline: `r crr_summary$years_over_one[crr_summary$scenario=="baseline"]`
* Years with CRR > 1 for alt: `r crr_summary$years_over_one[crr_summary$scenario=="alt"]`


## Life history diversity and fitness

### pHOS
```{r, out.width = '100%'}
phos <- baseline_results$phos |> 
  as.data.frame() |> 
  mutate(watershed = winterRunDSM::watershed_labels,
         scenario = "baseline") |> 
  bind_rows(alt_results$phos |> 
              as.data.frame() |> 
              mutate(watershed = winterRunDSM::watershed_labels,
                     scenario = "alt")) |> 
  pivot_longer(`1`:`20`,
               names_to = "sim_year",
               values_to = "phos") |> 
  filter(watershed %in% c("Upper Sacramento River")) |> 
  mutate(sim_year = as.numeric(sim_year)) 
  
ggplot(phos) + 
  geom_line(aes(sim_year, phos, color = scenario)) +
  geom_point(aes(sim_year, phos, color = scenario, shape = scenario), size = 3) +
  scale_color_manual(values = colors2)+
  theme_plots
```

### Percent years with fewer than 4 size classes represented in outmigration

### Highest proportion of smolts entering ocean in a 1-week period 

## Increase number of tribs supporting WR

### Number of adults spawning in tributaries

### Number of juveniles rearing in tributaries

### Number of independent populations

### Number of independent populations in historic habitats

### Number of dependent populations 