---
title: "Explore Intro Scenario"
author: "Cat Pien, Liz Stebbins, Flowwest"
format: html
editor: visual
---

# Example Portfolio Model Results

## Description of included actions

* H1: 
* H2b: 
* H2c: 
* H3:
* SR4a: 
* SR11:
* O2: 
* BC1:
* BC8: 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r message = FALSE, warning = FALSE}
library(winterRunDSM)
library(tidyverse)
library(plotly)
library(ggplot2)
library(zoo)
library(kableExtra)

theme_plots <- 
  theme_bw() + 
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 13),
        legend.position = "top")

colors <- c("#96C8A2", "#006A4E")
colors2 <- c("darkgoldenrod2", "coral4")
```


```{r, warning = FALSE}
# baseline
baseline_seeds <- winterRunDSM::winter_run_model(scenario = NULL, 
                                                 mode = "seed",
                                                 seeds = NULL, 
                                                 ..params = winterRunDSM::wr_sdm_baseline_params)

baseline_results <- winterRunDSM::winter_run_model(mode = "simulate", 
                                                   ..params = winterRunDSM::wr_sdm_baseline_params,
                                                   seeds = baseline_seeds)

# test out scenario 
portfolio_eg_seeds <- winterRunDSM::winter_run_model(scenario = NULL, 
                                            mode = "seed",
                                            seeds = NULL, 
                                            ..params = winterRunDSM::wr_sdm_scen_intro_params)

portfolio_eg_results <- winterRunDSM::winter_run_model(mode = "simulate", 
                                              ..params = winterRunDSM::wr_sdm_scen_intro_params,
                                              seeds = portfolio_eg_seeds)
```

## Summary Table of all the metrics 


## Abundance 

### Spawners

```{r, out.width = '100%'}
spawners_split <- baseline_results$spawners |> 
  as.data.frame() |> 
  mutate(watershed = winterRunDSM::watershed_labels,
         scenario = "baseline") |> 
  bind_rows(portfolio_eg_results$spawners |> 
              as.data.frame() |> 
              mutate(watershed = winterRunDSM::watershed_labels,
                     scenario = "portfolio_eg")) |> 
  pivot_longer(`1`:`20`,
               names_to = "sim_year",
               values_to = "spawners") |> 
  filter(watershed %in% c("Upper Sacramento River", "Battle Creek")) |> 
  mutate(sim_year = as.numeric(sim_year)) 

spawners <- spawners_split |> 
  group_by(sim_year, scenario) |> 
  summarize(spawners = sum(spawners))

mean_spawners <- spawners |> 
  group_by(scenario) |> 
  summarize(mean_spawners = round(mean(spawners)))

plotly::ggplotly(ggplot(spawners_split) + 
  geom_line(aes(sim_year, spawners, color = scenario)) +
  geom_point(aes(sim_year, spawners, color = scenario, shape = scenario), size = 3) +
    facet_wrap(~watershed, nrow = 2) + 
  scale_color_manual(values = colors)+
  theme_plots)
```


```{r}
plotly::ggplotly(ggplot(spawners) + 
  geom_line(aes(sim_year, spawners, color = scenario)) +
  geom_point(aes(sim_year, spawners, color = scenario, shape = scenario), size = 3) +
  scale_color_manual(values = colors)+
  theme_plots)
```


```{r}
kable(mean_spawners)
```


### Natural origin WR returning to spawn

Natural origin spawners - each year and tributary has a different proportion natural based on hatchery allocations 
Total combines applicable watersheds

```{r, echo = FALSE, out.width = '100%'}
spawn_prop <- portfolio_eg_results$proportion_natural_at_spawning[c("Upper Sacramento River", "Battle Creek"),] |> 
  as.data.frame() |> 
  mutate(watershed = c("Upper Sacramento River", "Battle Creek")) |> 
  pivot_longer(`1`:`20`,
               names_to = "sim_year",
               values_to = "prop_natural") |> 
  mutate(sim_year = as.numeric(sim_year))

returns_split <- left_join(spawners_split, spawn_prop) |> 
  mutate(nat_returns = spawners * prop_natural,
         hatchery_returns = spawners - nat_returns)

returns <- returns_split |> 
  group_by(scenario, sim_year) |> 
  summarize(nat_returns = sum(nat_returns),
            hatchery_returns = sum(hatchery_returns))

mean_natural_return <- returns |>
  group_by(scenario) |>
  summarize(mean_natural_return = round(mean(nat_returns))) 

ggplot(returns) +
  # geom_col(aes(return_sim_year, returns, fill = scenario), color = "black", position = position_dodge2()) +
  geom_line(aes(sim_year, nat_returns, color = scenario)) +
  geom_point(aes(sim_year, nat_returns, color = scenario, shape = scenario), size = 3) +
  labs(title = "Natural Origin Returning Adults") +
  # facet_wrap(~watershed, nrow = 2) + 
  scale_color_manual(values = colors)+
  theme_plots
```

```{r}
kable(mean_natural_return)
```

### Hatchery origin WR returning to spawn

Hatchery origin spawners - each year has a different proportion natural based on hatchery allocations 
Total combines applicable watersheds

```{r, echo = FALSE, out.width = '100%'}
mean_hatchery_return <- returns |>
  group_by(scenario) |>
  summarize(mean_hatchery_return = round(mean(hatchery_returns))) 

ggplot(returns) +
  # geom_col(aes(return_sim_year, returns, fill = scenario), color = "black", position = position_dodge2()) +
  geom_line(aes(sim_year, hatchery_returns, color = scenario)) +
  geom_point(aes(sim_year, hatchery_returns, color = scenario, shape = scenario), size = 3) +
  labs(title = "Hatchery Origin Returning Adults") +
  # facet_wrap(~watershed, nrow = 2) + 
  scale_color_manual(values = colors2)+
  theme_plots
```

```{r}
kable(mean_hatchery_return)
```

### Frequency of population decline

Number of years of decline in spawners

```{r}
decline <- spawners |> 
  group_by(scenario) |> 
  mutate(decline = if_else(spawners < lag(spawners,1), 1L, 0L)) |> 
  ungroup()

total_declines <- decline |> 
  group_by(scenario) |> 
  summarize(total_declines = sum(decline, na.rm = TRUE))
```


```{r}
kable(total_declines)
```


### Frequency of catastrophic decline

Greatest 3-year average change in spawners (escapement)

```{r}
change_spawn <- spawners |> 
  group_by(scenario) |> 
  mutate(change_abund = spawners-lag(spawners, 1),
         prop_change_abund = change_abund/lag(spawners,1)) |> 
  mutate(avg_change = rollapply(change_abund, width = 3, FUN = mean, align = "right", partial = FALSE, fill = NA),
         avg_prop_change = rollapply(prop_change_abund, width = 3, FUN = mean, align = "right", partial = FALSE, fill = NA) ) |> 
  ungroup()

max_declines <- change_spawn |> 
  group_by(scenario) |> 
  summarize(max_decline = round(min(avg_prop_change, na.rm = TRUE),3)) 
```

```{r}
kable(max_declines)
```


## Productivity

### Number of juveniles emigrating from upper Sacramento River

Number of juveniles coming out of Upper Sacramento River before rear, grow, migrate, moving to other watersheds

```{r}
juvs_us <- baseline_results$juveniles |>
  mutate( scenario = "baseline") |>
  bind_rows(portfolio_eg_results$juveniles |> mutate(scenario = "portfolio_eg")) |> 
  filter(watershed %in% c("Upper Sacramento River")) |> 
  group_by(watershed, scenario, year) |> 
  summarize(total_juv = sum(juveniles))
  
mean_juv_us <- juvs_us |> 
  group_by(scenario) |> 
  summarize(mean_juv = mean(total_juv))

ggplot(juvs_us) +
  # geom_col(aes(return_sim_year, returns, fill = scenario), color = "black", position = position_dodge2()) +
  geom_line(aes(year, total_juv, color = scenario)) +
  geom_point(aes(year, total_juv, color = scenario, shape = scenario), size = 3) +
  labs(title = "Juveniles in Upper Sac", y = "Juveniles in Upper Sac") +
  scale_color_manual(values = colors)+
  theme_plots
```

```{r}
kable(mean_juv_us)
```


### Number of natural origin smolts surviving to reach the ocean 

This metric calculates juveniles at Chipps using large and very large fish (smolts) originating from the Upper Sacramento River.

```{r, out.width = '100%'}
juv_chipps <- baseline_results$juveniles_at_chipps |> 
  as.data.frame() |> 
  mutate(scenario = "baseline") |> 
  bind_rows(portfolio_eg_results$juveniles_at_chipps |> 
              as.data.frame() |> 
              mutate(scenario = "portfolio_eg")) |>  
  filter(watershed %in% c("Upper Sacramento River"),
         size %in% c("l", "vl")) |> 
  group_by(watershed, scenario, year) |> 
  summarize(jac = sum(juveniles_at_chipps)) |> 
  ungroup()

mean_jac <- juv_chipps |> 
  group_by(scenario, watershed) |> 
  summarize(mean_jac = round(mean(jac)))

ggplot(juv_chipps) + 
  # geom_col(aes(return_sim_year, returns, fill = scenario), color = "black", position = position_dodge2()) +
  geom_line(aes(year, jac, color = scenario)) +
  geom_point(aes(year, jac, color = scenario, shape = scenario), size = 3) +
  labs(title = "Juveniles at Chipps", y = "Juveniles at Chipps") +
  scale_color_manual(values = colors)+
  theme_plots
```

```{r}
kable(mean_jac)
```


### Cohort Replacement Rate

Battle Creek does not have baseline so keeping both for now

```{r, out.width = '100%'}
crr <- spawners_split |> 
  group_by(scenario, watershed) |> 
  mutate(crr = spawners/lag(spawners, 3),
         crr_over_one = if_else(crr > 1, 1L, 0L)) |> 
  ungroup()

crr_summary <- crr |> 
  group_by(scenario, watershed) |> 
  summarize(years_over_one = sum(crr_over_one, na.rm = TRUE),
         mean_crr = round(mean(crr, na.rm = TRUE),3)) |> 
  ungroup() 

ggplot(crr) + 
  geom_line(aes(sim_year, crr, color = scenario)) +
  geom_point(aes(sim_year, crr, color = scenario, shape = scenario), size = 3) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  labs(title = "Cohort Replacement Rate", y = "Cohort Replacement Rate") +
  facet_wrap(~watershed, nrow = 2) + 
  scale_color_manual(values = colors)+
  theme_plots
```

```{r}
kable(crr_summary)
```


## Life history diversity and fitness
Add labels to the y axis 

### pHOS
```{r, out.width = '100%'}
phos <- baseline_results$phos |> 
  as.data.frame() |> 
  mutate(watershed = winterRunDSM::watershed_labels,
         scenario = "baseline") |> 
  bind_rows(portfolio_eg_results$phos |> 
              as.data.frame() |> 
              mutate(watershed = winterRunDSM::watershed_labels,
                     scenario = "portfolio_eg")) |> 
  pivot_longer(`1`:`20`,
               names_to = "sim_year",
               values_to = "phos") |> 
  filter(watershed %in% c("Upper Sacramento River", "Battle Creek")) |> 
  mutate(sim_year = as.numeric(sim_year)) 
  
ggplot(phos) + 
  geom_line(aes(sim_year, phos, color = scenario)) +
  geom_point(aes(sim_year, phos, color = scenario, shape = scenario), size = 3) +
  scale_color_manual(values = colors2)+
  labs(title = "pHOS") +
  facet_wrap(~watershed, nrow = 2) + 
  theme_plots
```

### Size class diversity

Shannon diversity measures evenness and diversity.
A higher number indicates greater evenness and diversity. A low number may indicate domination by one group (in this case, size class). 

```{r, out.width = '100%'}
juvenile_size_ocean_entry <- baseline_results$juveniles_at_chipps |>
    group_by(year, watershed, size_or_age = size) |>
    summarise(value = sum(juveniles_at_chipps, na.rm = TRUE)) |>
    mutate(scenario = "baseline") |> bind_rows(
  portfolio_eg_results$juveniles_at_chipps |>
    group_by(year, watershed, size_or_age = size) |>
    summarise(value = sum(juveniles_at_chipps, na.rm = TRUE)) |>
    mutate(scenario = "portfolio_eg")) |> 
  filter(watershed %in% c("Upper Sacramento River")) |> 
  mutate(size_or_age = factor(size_or_age, levels = c("s", "m", "l", "vl")))

  annual_total <- juvenile_size_ocean_entry |>
    group_by(year, scenario) |>
    summarize(total_juveniles = sum(value, na.rm = T))

  shannon_di_size <- juvenile_size_ocean_entry |>
    group_by(year, size_or_age, scenario) |>
    summarize(frequency = sum(value, na.rm = T)) |>
    ungroup() |>
    left_join(annual_total) |>
    mutate(pi = frequency / total_juveniles,
           ln_pi = log(pi),
           pi_ln_pi = pi * ln_pi) |>
    group_by(year, scenario) |>
    summarize(shannon_index = -1 * sum(pi_ln_pi, na.rm = T)) |>
    ungroup()
  
  mean_shannon_di_size <- shannon_di_size |>
    group_by(scenario) |>
    summarise(avg_annual_shannon_di = round(mean(shannon_index),2)) |> 
    ungroup()
  
  ggplot(shannon_di_size) + 
  geom_line(aes(year, shannon_index, color = scenario)) +
  geom_point(aes(year, shannon_index, color = scenario, shape = scenario), size = 3) +
  labs(title = "Size Diversity")+
  scale_color_manual(values = colors)+
  theme_plots
```

```{r, out.width = '100%'}
ggplot(juvenile_size_ocean_entry)+
  geom_col(aes(year, value, fill = factor(size_or_age)), position = "fill", color = "black") +
  labs(y = "proportion of juveniles", title = "Size Diversity") +
  facet_wrap(~scenario, nrow = 2) + 
  viridis::scale_fill_viridis(option= "mako", discrete = TRUE)  + 
  theme_bw()
```

```{r}
kable(mean_shannon_di_size)
```


### Timing diversity 

Shannon diversity measures evenness and diversity.
A higher number indicates greater evenness and diversity. A low number may indicate domination by one group (in this case, month of ocean entry).

```{r, out.width = '100%'}
juvenile_month_ocean_entry <- baseline_results$juveniles_at_chipps |>
    group_by(year, watershed, month) |>
    summarise(value = sum(juveniles_at_chipps, na.rm = TRUE)) |>
    mutate(scenario = "baseline") |> bind_rows(
  portfolio_eg_results$juveniles_at_chipps |>
    group_by(year, watershed, month) |>
    summarise(value = sum(juveniles_at_chipps, na.rm = TRUE)) |>
    mutate(scenario = "portfolio_eg")) |> 
  filter(watershed %in% c("Upper Sacramento River"))

  annual_total <- juvenile_month_ocean_entry |>
    group_by(year, scenario) |>
    summarize(total_juveniles = sum(value, na.rm = T))

  shannon_di_timing <- juvenile_month_ocean_entry |>
    group_by(year, month, scenario) |>
    summarize(frequency = sum(value, na.rm = T)) |>
    ungroup() |>
    left_join(annual_total) |>
    mutate(pi = frequency / total_juveniles,
           ln_pi = log(pi),
           pi_ln_pi = pi * ln_pi) |>
    group_by(year, scenario) |>
    summarize(shannon_index = -1 * sum(pi_ln_pi, na.rm = T)) |>
    ungroup()
  
  mean_shannon_di_timing <- shannon_di_timing |>
    group_by(scenario) |>
    summarise(avg_annual_shannon_di = round(mean(shannon_index),2)) |> 
    ungroup() 
    
  
  ggplot(shannon_di_timing) + 
  geom_line(aes(year, shannon_index, color = scenario)) +
  geom_point(aes(year, shannon_index, color = scenario, shape = scenario), size = 3) +
  labs(title = "Timing Diversity")+
  scale_color_manual(values = colors)+
  theme_plots
```
```{r, out.width = '100%'}
ggplot(juvenile_month_ocean_entry)+
  geom_col(aes(year, value, fill = factor(month)), position = "fill", color = "black") +
  labs(y = "proportion of juveniles", title = "Timing Diversity") +
  facet_wrap(~scenario, nrow = 2) + 
  viridis::scale_fill_viridis(option= "turbo", discrete = TRUE)  + 
  theme_bw()
```

```{r}
kable(mean_shannon_di_timing)
```


### Combined size-timing diversity 

Shannon diversity measures evenness and diversity.
A higher number indicates greater evenness and diversity. A low number may indicate domination by one group (in this case, size and month combination groupings).

```{r, out.width = '100%'}
juvenile_size_timing_ocean_entry <- baseline_results$juveniles_at_chipps |>
    group_by(year, month, watershed, size_or_age = size) |>
    summarise(value = sum(juveniles_at_chipps, na.rm = TRUE)) |>
    mutate(scenario = "baseline") |> bind_rows(
  portfolio_eg_results$juveniles_at_chipps |>
    group_by(year, month, watershed, size_or_age = size) |>
    summarise(value = sum(juveniles_at_chipps, na.rm = TRUE)) |>
    mutate(scenario = "portfolio_eg")) |> 
  filter(watershed %in% c("Upper Sacramento River", "Battle Creek"))

annual_total <- juvenile_size_timing_ocean_entry |>
    group_by(year, scenario) |>
    summarize(total_juveniles = sum(value, na.rm = T))
  
shannon_di_size_timing <- juvenile_size_timing_ocean_entry |>
    group_by(year, size_or_age, scenario, month) |>
    summarize(frequency = sum(value, na.rm = T)) |>
    ungroup() |>
    left_join(annual_total) |>
    mutate(pi = frequency / total_juveniles,
           ln_pi = log(pi),
           pi_ln_pi = pi * ln_pi) |>
    group_by(year, scenario) |>
    summarize(shannon_index = -1 * sum(pi_ln_pi, na.rm = T)) |>
    ungroup()
  
mean_shannon_di_size_timing <- shannon_di_size_timing |>
    group_by(scenario) |>
    summarise(avg_annual_shannon_di = round(mean(shannon_index),2))
```

```{r}
kable(mean_shannon_di_size_timing)
```

```{r, out.width = '100%'}
ggplot(shannon_di_size_timing) + 
  geom_line(aes(year, shannon_index, color = scenario)) +
  geom_point(aes(year, shannon_index, color = scenario, shape = scenario), size = 3) +
  scale_color_manual(values = colors2)+
  labs(title = "Shannon Diversity for Timing and Size") +
  theme_plots
  
```

## Increase number of tribs supporting WR

### Number of adults spawning in tributaries

In this case, just Battle Creek

```{r}
spawners_in_tribs <- spawners_split  |>  filter(watershed == "Battle Creek")

mean_spawners_in_tribs <- spawners_in_tribs |> 
  group_by(scenario) |> 
  summarize(mean_spawners = round(mean(spawners)))

ggplot(spawners_in_tribs) + 
  geom_line(aes(sim_year, spawners, color = scenario)) +
  geom_point(aes(sim_year, spawners, color = scenario, shape = scenario), size = 3) +
  labs(title = "Spawners in Battle Creek")+
  scale_color_manual(values = colors)+
  theme_plots
```

```{r}
kable(mean_spawners_in_tribs)
```

### Number of juveniles rearing in tributaries

This tracks juveniles in the tributary, right before they grow and rear. 

```{r, out.width = '100%'}
juvs_trib <- baseline_results$juveniles |>
  mutate(scenario = "baseline") |>
  bind_rows(portfolio_eg_results$juveniles |> mutate(scenario = "portfolio_eg")) |> 
  filter(watershed %in% c("Battle Creek")) |> 
  group_by(watershed, scenario, year) |> 
  summarize(total_juv = sum(juveniles))
  
mean_juv_trib <- juvs_trib |> 
  group_by(scenario) |> 
  summarize(mean_juv = mean(total_juv))

ggplot(juvs_trib) +
  # geom_col(aes(return_sim_year, returns, fill = scenario), color = "black", position = position_dodge2()) +
  geom_line(aes(year, total_juv, color = scenario)) +
  geom_point(aes(year, total_juv, color = scenario, shape = scenario), size = 3) +
  labs(title = "Juveniles in Battle Creek", y = "Juveniles in Battle Creek") +
  scale_color_manual(values = colors)+
  theme_plots
```

```{r}
kable(mean_juv_trib)
```


### Number of independent populations

Independent populations, also named viable populations, include those with
* greater than 500 spawners
* pHOS less than 5%
* CRR > 1
* Growth rate (spawner change) > 1

```{r, out.width = '100%'}
ind_pop <- spawners |> left_join(phos) |> left_join(crr) |> left_join(change_spawn) |> 
  mutate(growth_rate = change_abund/lag(spawners)) |> 
  mutate(above_500_spawners = if_else(spawners > 500, 1L, 0L),
         phos_less_than_5_percent = if_else(phos < .05, 1L, 0L),
         crr_above_1 = if_else(crr >= 1, 1L, 0L),
         growth_rate_above_1 = if_else(growth_rate >= 0, 1L, 0L),
         independent_population = if_else(above_500_spawners & phos_less_than_5_percent &
                                           growth_rate_above_1 & crr_above_1, 1L, 0L))

ind_pop_long <- ind_pop |> 
  pivot_longer(cols = c(above_500_spawners, phos_less_than_5_percent, crr_above_1, growth_rate_above_1), names_to = "metric", values_to= "value")

ind_pop_summary <- ind_pop |> 
  group_by(scenario, watershed) |> 
  summarize(independence_years = sum(independent_population)) |> 
  arrange(watershed)
```

```{r, out.width = '100%'}
ggplot(ind_pop_long) + 
  geom_tile(aes(sim_year, metric, fill = value), color = "black")+
  facet_grid(watershed~scenario)
```
```{r}
kable(ind_pop_summary)
```

### Number of independent populations in historic habitats

- Not relevant for this portfolio since only Battle Creek included. Historic habitats include Pit River, Little Sacramento River, and McCloud River. 

### Number of dependent populations 

Defined as populations not meeting independent population criteria

- 2 populations 
